---
title: "Open software and stuff"
subtitle: "Free yourself &#x2694;&#xFE0F; &#x1F60E; Be happy"  
author: 
  - "Timothée Bonnet"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: 
      - default
      - default-fonts
      - xaringan-themer.css
      - "https://cdnjs.cloudflare.com/ajax/libs/animate.css/3.7.0/animate.min.css"
    lib_dir: libs
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
      countdown: 60000
---
class: center, middle

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  fig.show = TRUE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "#FF961C",
  inverse_header_color = "#FFFFFF"
)
```


## 1. Operating System &#x1F427;

## 2. Why open-source / free software? 

## 3. Science stuff &#x1F52C;

## 4. Normal life stuff &#x1F63A;



---
class: center, middle, animated, slideInRight

# Linux (Ubuntu): What you may think it is

<video width="768" height="576" controls>
  <source src="media/vim.mp4" type="video/mp4">
</video>


---
class: inverse, center, middle, animated, bounceInDown

# Linux (Ubuntu): What it really is

<video width="768" height="576" controls>
  <source src="media/Rstudiowork.mp4" type="video/mp4">
</video>


---

## Linux (Ubuntu): What it really is

> Command line powerful but unecessary
>
> Installation super easy
>
> Start with double-boot (and soon never use Windows/MacOS again)


### Compared to Windows

### Compared to MacOS

---
class: center, middle

## I am no expert but... 
### Proprietary software slows down science

--

### ... prevents human progress

--

### Is a source of broad inequality


---
class: middle, animated, bounceInDown

# Data management, analysis, statistics...

--

### Supports science that can be shared, checked, reproduced, improved

Who can pay for licenses out of first world universities?

How do you check analyses that were run by clicking in menus? 

How do you contribute new/improved features if you don't work for the owner?

--

## A few examples:
### Statistics, data: SPSS/SAS/Matlab  &#8594; R/Python/Scilab...
### Workflows: XXX  &#8594; Make, Conda, Git


---
class: middle, animated, bounceInDown


# Academic office work

* Slides: Powerpoint  &#8594; LibreOffice Impress, $\LaTeX$ Beamer, RPres, Markdown Xaringan <img src="https://upload.wikimedia.org/wikipedia/commons/b/be/Sharingan_triple.svg" height="40px" />
* Writing: Microsoft Word &#8594; LibreOffice, $\LaTeX$, Markdown, Pandoc...
* Spreadsheet: Microsoft Excel &#8594; LibreOffice Calc
* Reference manager: <b>Mendeley</b>, Endnote &#8594; Zotero


---
class: middle, animated, bounceInDown

name: colors

# Open/free software not only for science!
## Basic officework software

* OS: MacOS / Windows &#8594; Linux (Ubuntu, Debian...)
* Browsing: Chrome, Edge... &#8594; Firefox, Chromium...
* Emails: ?? &#8594; Thunderbird, Evolution...
* Music: ??  &#8594; foobar

## Specialised software

* Audio: Audacity
* Photo: Darktable, RawTherapy
* Video, 3D, video game: Blender

---

### Blockquote

> This is a blockquote following a header.
>
> When something is important enough, you do it even if the odds are not in your favor.

---

### Code Blocks

#### R Code

```{r eval=FALSE}
ggplot(gapminder) +
  aes(x = gdpPercap, y = lifeExp, size = pop, color = country) +
  geom_point() +
  facet_wrap(~year)
```

#### JavaScript

```js
var fun = function lang(l) {
  dateformat.i18n = require('./lang/' + l)
  return true;
}
```



---

```{r finalslide, include=FALSE}
options(htmltools.dir.version = FALSE)
```
class: centre, animated, flipInY
background-image: url(https://upload.wikimedia.org/wikipedia/commons/b/be/Sharingan_triple.svg)

## BTW, Xaringan is awesome! 
.large[(yes, what you see is R-Markdown in RStudio)]

<br> 
<br> 
<br> 
<br> 
<br> 
<br> 
<br> 
<br>
<br> 


[**https://github.com/yihui/xaringan**](https://github.com/yihui/xaringan)<br>
[https://github.com/gadenbuie/xaringanthemer](https://github.com/gadenbuie/xaringanthemer)

The chakra comes from [remark.js](https://remarkjs.com), [**knitr**](http://yihui.name/knitr), and [R Markdown](https://rmarkdown.rstudio.com). Image credit: [Wikimedia Commons](https://commons.wikimedia.org/wiki/File:Sharingan_triple.svg)
